Theory about Normal Basis on $F(2^k)$

### 2016-12-08 dissertation/intro.tex
dissertation v2:Appendix update 1
dissertation v2:reconstructing part intro

\chapter{Introduction}
\label{ch:intro}
With the increasing size of integrated
circuits, sequential circuit designers face complicated problems of
design errors in specification models and implementations. These
errors are usually modeled as ``bad" states, and the
circuits/functional components are modeled as finite state machines
(FSMs). Once state reachability is analyzed, the existence of errors
can be identified by checking whether ``bad" states are {\it
  reachable} from certain initial states. Temporal logic {\it model
  checking} (MC) formulations and solvers are often used for this
purpose. Once the designs and specification models are validated using
model checking, optimized implementations of sequential circuits are
synthesized. A subsequent problem then needs to be solved to prove
that the sequential circuit implementations are equivalent to the
original specification models, i.e. {\it Sequential Equivalence
  Checking   (SEC)}. When the specification is given as an arithmetic function
which canonically represents the circuit, then the problem 
becomes {\it functional verification} of sequential arithmetic circuits.

@@

For word-level SEC, we are given two designs, or their corresponding
Mealy/Moore FSMs ${\mathcal{M}}_1,{\mathcal{M}}_2$, along with their
initial starting states $S_0^1,S_0^2$. We wish to prove the absence of
a sequence of inputs (string) that distinguishes the initial
states \cite{coudert:iccad90}
\cite{coudert1990verification}. Fundamentally, this requires 
the construction of a product machine; and the main research problem
relates to that of performing {\it FSM traversal} \cite{touati1990implicit}
but  at the word-level. Analogously, in the case of MC, the problem
is setup w.r.t. a FSM $\mathcal{M}$, a set of initial states $S_0$ and
a set of property states $p$. Techniques are to be researched that
verify that there exist no sequence of transitions from an initial
state to a non-property state (``bad'' state). These problems have to
be solved in the context of word-level verification -- i.e. data
representation, abstraction using UNSAT cores and
algorithm execution has to be carried out at word-level.

\section{Hardware Design and Verification Overview}
Following the Moore's law, the level of integration on modern VLSI becomes very high, 
such that a single digital VLSI design can implement a true system-level component.
The design process also evolves from manual design with little validation, to 
a formal 3-step procudure which needs collaboration of team with large number of 
engineers. The 3 major steps are: 1) Design, which is to specify and enter the design intent;
2) Implement, which is to refine the design through all phases with the assistance of Computer-Aided-Design (CAD)
tools; 3) Verify, which is to verify the correctness of design and implementation.

Nowadays the verify step is usually completed by a team with specialities on test, verification and validation of 
circuits. This step is also automated as an indispensable part of CAD tools, when circuit synthesis is performed. 
Figure \ref{fig:designflow} shows the typical synthesis flow, which covers all procudures starting from 
register-transfer-level (RTL) description (using hardware design languages, i.e. HDL) to  the 
physical design on silicon (depicted by the layout). The objective of verification in synthesis is 
to ensure the implementation is consistent with the original design intent. Verification is 
an important quality control measure before sending design layout to the VLSI foundaries.
Considering the high cost of fabrication, any faults and errors in the design will bring considerable 
waste of funds for the designers. On the other hand, all aspects of the society increasingly depend on 
the stability and accuracy of digital VLSI circuits, even small flaws or short-time failures can cause 
huge loss, especially in medical applications, military facilities and financial systems.
Therefore, it is of utmost importance to verify the correctness of VLSI designs.

One way to perform the verification is called {\it simulation}. It is defined as the collage of all circuit validation 
methods which apply stimulations on the inputs of circuit model and verify correctness of the outputs.
However, simulation is not a complete solution to circuit verification problems. In modern designs with 
large number of logic components and complicated architecures, it is impractical to simulate all possible 
test vectors. Usually only test vectors that correspond to typical failures are selected in simulation, which 
cannot cover unexpected failure patterns caused by special inputs. The notorious Intel FDIV bug \cite{nicely:FDIV}
is a good example when simulation fails. Failure occurs when only 5 entries in an $84\times 16$ look-up table
(LUT) are activated,  which is rarely used in most divisions. Because of the limitation of simulation,  
test engineers from Intel failed to detect the bug,  which brought $\$475$ million dollors recalling bill
for the company. Thus,  new methods that can guarantee the correctness of the design are necessary to be explored.

\section{Formal Verification}
% TODO this part require less modifications
Another method developed besides simulation is \emph{formal verification}, it utilizes 
mathematical theory to reason about the correctness of hardware designs.
Formal verification can provide $100\%$ fault coverage from 2 aspects. On the one hand,  it formalize properties
for the circuit model which are irrelevant to specific input signals,  and prove them mathematically.
On the other hand,  it adopts formal languages to strictly describe the design intents and detailed implementation, 
and deduces circuit function from the implementation. These descriptions are named as {\it specifications}.
Formal verification has two main forms: property checking and equivalence 
checking.

{\it Property checking} (or property verification) verifies
that a design satisfies certain given properties. Property checking is done mainly 
in the form of theorem proving, model checking, or approaches which 
combine the two.
\begin{enumerate}
\item \emph{Theorem proving} \cite{theoremproving:91} requires the existence of
mathematical descriptors of the specification and implementation of the 
circuit. Theorem provers apply mathematical rules to these descriptors to
derive new properties of the specification. In this way, the tool can reduce
a proof goal to simpler sub-goals, which can be automatically verified.
However, generating the initial proof-goal requires extensive guidance from
the user, so there is an overall lack of automation in theorem 
proving.
\item \emph{Model checking} \cite{modelcheck:99} is an approach
to verifying finite-state systems where specification 
properties are modeled as a system of 
logic formulas. The design is then traversed to check if the 
properties hold. If the design is found to violate a
particular property, a counter-example is generated which exercises the
incorrect behavior in the design. Such counter-examples allow the designer
to trace the behavior and find where the error in the design lies.
Modern model checking techniques use the result to automatically refine
the system and perform further checking.
These tools are typically automated, and thus have found widespread 
use in CAD tool suites.
\end{enumerate}

{\it Equivalence Checking} verifies that two different representations of
a circuit design have equivalent functionality. An example of equivalence
checking as it applies to the hardware design flow is shown in
Fig. \ref{fig:equivflow}.

{
\begin{figure}[h]
\centerline{
\includegraphics[width=0.5\textwidth]{./figures/designVerification}
}
\caption{Equivalence checking as applied to the hardware design flow.}
\label{fig:equivflow}
\end{figure}
}

There are two major
equivalence checking techniques: graph-based
and satisfiability-based.
\begin{enumerate}
\item \emph{Graph-based} techniques construct a canonical graph 
representation, such as a Binary Decision Diagram (\emph{BDD}) or one of
its many variants, of each circuit. A linear comparison is then conducted to 
determine whether the two graphs are isomorphic. Since the graph 
representation is canonical, the graphs of the two circuits will be 
equivalent if and only if the circuits perform the same function.
\item \emph{Satisfiability} techniques construct a miter of the two circuits,
typically in a graph such as an And-Inverter graph (\emph{AIG}). A
\emph{miter} is a combination of the two circuits with one bit-level output, which 
is only in a "1" state when the outputs of the circuits differ given 
the same given 
input, as shown in Fig. \ref{fig:miter}. 
A satisfiability (\emph{SAT}) tool \cite{csat} 
is then employed to simplify the graph and find a solution to the miter, 
i.e. find an input for which the 
miter output is "1". If a solution is found, this solution acts as a 
counter-example of when the circuit outputs differ; otherwise the circuits
are functionally equivalent.
\end{enumerate}

{
\begin{figure}[h]
\centerline{
\includegraphics[width=0.8\textwidth]{./figures/betterMiter}
}
\caption{A miter of two circuits.}
\label{fig:miter}
\end{figure}
}

%\subsection{Computer-Algebra Based Formal Verification}
Certain formal verification methods use \emph{computer-algebra} and \emph{algebraic
geometry} techniques based on mathematical theories.
Unlike SAT-based verification, modern algebraic geometry 
techniques do not explicitly solve the constraints to find a solution; 
rather, they reason about the presence or absence of solutions, or explore
the geometry of the solutions.
These methods \cite{Avrunin:CAV} \cite{condrat-tacas07} \cite{gbverify:2007} 
transform the circuit design into a polynomial system. Typically, this system
of polynomials is then used to compute a Gr\"obner basis \cite{gb_book}. 
Computation of Gr\"obner bases allows for 
the easy deduction of important properties of a polynomial system, 
such as the presence or absence of 
solutions. These properties are then leveraged to perform 
verification. Unfortunately, such a computation 
has been shown to be doubly 
exponential in the worst case, and thus these methods have not been 
practical for real-world applications. However, recent
breakthroughs in computer-algebra hardware verification have shown
that it is possible to overcome the complexity of this computation while
still utilizing the beneficial properties of a Gr\"obner bases
\cite{lv:phd}.

Word-level abstractions have a wide variety applications in formal 
verification. Theorem proving techniques can leverage abstraction as an 
automatic decision procedure or as a canonical reduction 
engine. For example, since RTL is composed of circuit blocks that represent 
the underlying circuit, RTL verification methods can exploit 
abstractions of these blocks.
This is seen in the following RTL verification methods:
\begin{itemize}  
\item Model checking \cite{kroening:model}, 
where an approximation abstraction of RTL blocks is generated and then 
refined.
\item Graph-based equivalence 
checking \cite{WLS} \cite{arditi:bmd}, where abstraction methods are used
to generate a canonical word-level graph representation of the circuit.
\item Satisfiability-based equivalence checking \cite{lpsat}, where 
abstractions are used identify symmetrics and similarities in order to 
minimize the amount of logic that is sent to the 
SAT tool. 
\end{itemize}

Other equivalence checking techniques that employ abstractions 
include satisfiability modulo theory ({\it SMT}) techniques \cite{boolector} \cite{bryant:tacas07}, 
which are similar to SAT except they operate on higher-level data
structures (integers, reals, bit vectors, etc.), as well as 
constraint solving techniques \cite {ms:research} \cite{tew:iccad08}.
In general, RTL equivalence checking approaches would ideally maintain a 
high-level of abstraction while still retaining sufficient lower-level 
functional details  (such as bit-vector size, precision, etc) 
\cite{gupta_survey}.

Word-level hardware abstractions also have applications in RTL and datapath 
synthesis \cite{demicheli:iccad_98} \cite{demicheli:dac_99}
\cite{demicheli:tcad_03}. 
Abstractions of circuits allow for design reuse, which allows for tool-automated 
synthesis of larger circuit blocks.
Since hardware design specifications tend to be word-level, synthesis tools 
can use these larger circuit blocks to generate and optimize the
datapaths and create the RTL of the system. Thus, in order for a circuit to 
be used by these automated synthesis tools, its word-level abstraction must
be known.

@@

\section{Objective, Motivation and Contribution of this Dissertation}
This research proposes a
set of new, promising approaches for {\it word-level representation,
reachability analysis and abstraction} with applications to SEC, sequential GF
arithmetic functional verification and 
$k$-BMC with abstraction refinement. Our techniques operate at the word-level and they are based
largely on the concepts from {\it algebraic geometry}.

@@

The use of algebraic geometry
for sequential circuit verification and symbolic model checking has
been presented before. Avrunin presented the
concept of symbolic MC using algebraic geometry in
\cite{Avrunin:CAV}. Later, in \cite{vardi-iasted07}, Vardi presented
GB-algorithms for CTL, LTL and 
bounded MC over Boolean rings. However, these approaches are a
straight-forward transformation of the problem to {\it bit-level}
Boolean GB engines which are used in lieu of BDDs or SAT solvers. All
the concepts of word-level reachability, abstraction-refinement using
interpolation or UNSAT cores, etc., that we desire was not the focus of
\cite{Avrunin:CAV} \cite{vardi-iasted07}.

{\bf The Algebraic Geometry Model and Rationale:} We will represent
the FSMs -- the transition relations -- by means of a set of
multi-variate polynomials with coefficients from the finite (Galois)
field $\Fkk$ of $2^k$ elements, i.e. polynomials in the ring
$\Fkk[x_1,\dots,x_d]$. Each state of a FSM is identified with a
Boolean assignment to a set of $k$-bit state register variables
$S=\{s_0,\dots,s_{k-1}\}$. Therefore, we can consider each ($k$-bit)
state as a word-level element $S$ of the finite field
$\Fkk$. Algorithms can directly operate on polynomials in word-level
variable $S$.

Boolean functions with $k$-bit inputs and $k$-bit outputs 
$f: \B^k \rightarrow \B^k,\B = \{0, 1\}$ can be construed as functions
$f: \Fkk \rightarrow \Fkk$. It is well-known that over the finite
field ($\Fq$) of  $q$ elements, {\it every function} $f: \Fq
\rightarrow \Fq$ is a polynomial function \cite{ff:1997}. Moreover,
there exists a unique canonical polynomial $\Func$ that describes $f$.
This implies that one can derive a {\it  canonical, polynomial
  abstraction} of the function as $Z = \Func(A)$ where $Z, A$ are
word-level symbols representing $k$-bit operands. The concept also
generalizes to functions with different input/output bit-vector sizes,
i.e. functions of the type $f: \B^n \rightarrow \B^m$, modeled as
polynomials over $f:{\mathbb{F}}_{2^k} \rightarrow
{\mathbb{F}}_{2^k}$, where $k=LCM(n,m)$ \cite{ff:1997}.  
This implies that the FSM's transition relations can be
represented as polynomial functions (ideals) in $\Fkk$, and values of
state variables can be represented as solutions to these polynomials
(variety of the ideal). Subsequently, the ideal-variety correspondences
in algebraic geometry can be applied to implement symbolic reasoning
about state reachability. Moreover, as ${\mathbb{F}}_2 \subset \Fkk$,
our model provides a single, unified and bit-precise representation
for both bit-level (${\mathbb{F}}_2$) and word-level ($\Fkk$)
constraints.

The decision and abstraction procedures in our setting will rely on
the theory and technology of {\it \Grobner bases (GB)}. GB-based
algebraic reasoning is very powerful; in fact it is known to be
strictly stronger than resolution \cite{CEI:stoc-96}. Therefore, in
light of the above discussion, using concepts from algebraic geometry
and \Grobner bases over $\Fkk$, we can introduce another dimension of
word-level abstraction to the techniques in sequential verification.

\subsection{Dissertation Contributions}
% TODO revise this to contributions
In this dissertation, we focus on sequential circuit verification problems and 
propose solutions consisting of three main contributions: 
\begin{enumerate}[{1)}]
\item A method to perform reachability analysis at the word-level. 
	The given FSM is modeled as a system of polynomials over finite field,
	where the state space is mapped to its solution space.
	Our proposed algorithm utilize concepts in algebraic geometry including ideals and varieties.
	It also forms the foundation for word-level SEC and MC by enabling word-level abstraction of the
  state-space.
\item Using the theory of FSM traversal, we apply the algorithm to verify the 
function of sequential GF arithmetic circuits. Our proposed approach uses GB-engines efficiently 
and can verify sequential multipliers with 162-bit datapaths.
\item We explore the concept, and the computation, of unsatisfiable (UNSAT)
  cores of  a set of polynomials using Gr\"obner bases calculation. We apply the UNSAT core extraction 
  to abstraction-refinement
  techniques such as bounded model checking (BMC). 
\end{enumerate}

\section{Dissertation Organization}
The rest of the dissertation is organized as
follows: Chapter \ref{ch:prev} reviews previous work, and analyzes their drawbacks with respect to 
the word-level sequential verification problem.
Chapter \ref{ch:prelim_GF} covers preliminary
concepts and notation on finite fields, and the methodology about design of multipliers in finite fields.
Chapter \ref{ch:ideals} provides a theoretical background of algebraic geometry and Gr\"obner bases.
Chapter \ref{ch:reacha} describes the basic
concept of word-level FSM traversal and introduces our proposed word-level FSM traversal algorithm. 
Chapter \ref{ch:normal} explores the application of FSM traversal algorithm on 
functional verification of sequential GF arithmetic circuits. Chapter \ref{ch:UNSAT} describes
algorithmic techniques to derive UNSAT cores of polynomial
ideals. It also demonstrates with the help of an example how abstraction via
UNSAT cores in algebraic geometry can simplify BMC. 
Chapter \ref{ch:conclude} concludes the dissertation and outlines potential future research for
continuation of this work.

>>>>

\chapter{Introduction}
\label{ch:intro}

\section{Hardware Design and Verification Overview}
Because of the Moore's law, the level of integration in modern VLSI systems is becoming higher and higher. 
As a result, an entire system with billions of transistors can be built upon a single chip.
The design process also evolves from manual design with little validation, to 
a formal 3-step procudure which needs collaboration of teams with large number of 
engineers. The 3 major steps are: 1) Design, which is to specify and enter the design intent;
2) Implement, which is to refine the design through all phases with the assistance of Computer-Aided-Design (CAD)
tools; 3) Verify, which is to verify the correctness of design and implementation.

Nowadays the verification step is usually completed by a team that specializes on test, verification and validation of 
circuits. This step is also automated as an indispensable part of CAD tools, when circuit synthesis is performed. 
Figure \ref{fig:designflow} shows the typical synthesis flow, which covers all procedures starting from 
register-transfer-level (RTL) description (using hardware design languages, i.e. HDL) to  the 
physical design on silicon (depicted by the layout). The objective of verification in synthesis is 
to ensure the implementation is consistent with the original design intent. Verification is 
an important quality control measure before sending design layout to the VLSI foundaries.
Considering the high cost of fabrication, any faults and errors in the design will bring considerable 
waste of funds for the designers. On the other hand, all aspects of the society increasingly depend on 
the stability and accuracy of digital VLSI circuits; even small flaws or short-time failures can cause 
huge loss, especially in medical applications, military facilities and financial systems.
Therefore, it is of utmost importance to verify the correctness of VLSI designs.

One way to perform verification is by {\it simulation}. It is defined as the collage of all circuit validation 
methods which apply stimulations on the inputs of circuit model and verify correctness of the outputs.
However, simulation is not a complete solution to circuit verification problems. In modern designs with 
large number of logic components and complicated architecures, it is impractical to simulate all possible 
test vectors. Usually only test vectors that correspond to typical failures are selected in simulation, which 
cannot cover unexpected failure patterns caused by special inputs. The notorious Intel FDIV bug \cite{nicely:FDIV}
is a good example where simulation failed. Failure occurs with input assignments which are rarely used in most divisions. Because of the limitation of simulation,  
test engineers from Intel failed to detect the bug,  which brought about $\$475$ million dollors recalling bill
for the company. Thus,  new methods that can guarantee the correctness of the design are necessary to be explored.

Another method developed besides simulation is \emph{formal verification}, it utilizes 
mathematical theory to reason about the correctness of hardware designs.
Formal verification can provide $100\%$ fault coverage from two aspects. On the one hand,  
it adopts formal languages to strictly describe the design intent and detailed implementation, 
and deduces circuit function from the implementation.
On the other hand,  it formalizes properties
for the circuit model which are relevant only to specific input signals,  and prove the properties mathematically. 
These descriptions are named as {\it specifications}.
Formal verification has two main forms: property checking and equivalence 
checking.

\section{Formal Verification: Property and Equilavence Checking}

{\it Property checking} (or property verification) verifies
that a design satisfies certain given properties. Property checking is done mainly 
in the form of theorem proving (TP), model checking (MC), or TP/MC hybrid approaches.
\begin{enumerate}[{1)}]
\item \emph{Theorem proving} \cite{theoremproving:91} 
is a method of reasoning and mathematical logic dealing with proving 
mathematical theorems. In the application to circuit property checking, 
the specification as well as the circuit implementation
is described as theorems in mathematical logic. Subsequently logic rules
are employed to deduce new objective theorem. In practice, the tool can reduce
a proof goal to simpler sub-goals for automatic verification.

\item \emph{Model checking} \cite{modelcheck:99} is a technique 
for verifying if the specification properties are violated in finite-state systems. In circuit verification 
domain, both the specification and the circuit implementation are modeled as a system of 
logic formulas. The finite-state system is then traversed to check if the 
properties are violated. If violation occurs, 
a counter-example is then generated as a transition firing path corresponds to the
false behavior in the design. 
Modern model checking techniques use the error-trace to automatically refine
the system and perform further checkings.
\end{enumerate}

{\it Equivalence Checking} verifies that two different representations of
a circuit design have equivalent functionality. It can be applied to 
multiple steps in the hardware design flow in Figure \ref{fig:designflow},
such as checking functional equivalence between HDL description and RTL,
checking RTL equivalence between RTL and synthesized/optimized netlist
,and checking layout verification between netlist and layout for fabrication.

There are three major
equivalence checking techniques: graph-based,
satisfiability-based (SAT-based) and induction-based.
\begin{enumerate}[{1)}]
\item \emph{Graph-based} techniques compare between 2 circuit implementations 
by representing them using canonical graphs. 
The earliest invented canonical graph is the \emph{Binary Decision Diagram} (BDD) \cite{BRYA86}.
Many variants branch out from BDD, some widely used variants include 
ZDD \cite{minato1993zero}, BMD \cite{bmd}, FDD \cite{okfdd}, etc. The comparison algorithms can 
determine whether the two graphs are isomorphic. The canonity of the graph representation guarantees that
the graphs correspond to the two circuits will be 
equivalent if and only if the circuits are implementing the same function.
\item \emph{Satisfiability} (SAT) techniques utilize the satisfiability theory.
In circuit equivalence checking, a miter of the two circuits is created.
A \emph{miter} is a combination of the two circuits with one bit-level output, which 
gives output ``1" only when the outputs of the circuits differ with 
the same inputs, {\it e.g. }inputs $a,b,c$ shown in Figure \ref{fig:miter}. 
A SAT solver \cite{csat,mishchenko2006improvements} is then employed to simplify the problem 
and find a satisfying assignment to the inputs for which the 
miter output is ``1". If such an assignment exists, this solution acts as a 
counter-example of when the circuit outputs differ; otherwise the circuits
are functionally equivalent.
\item \emph{Induction-based} techniques are developed and applied to verify the 
equivalence between sequential circuits, which is called the {\it sequential equivalence checking}
(SEC) problem. A miter model can also be built with two sequential circuits. Through the 
miter model, SEC problem is transformed into a sequential backward justification problem.
Equivalence of states and transitions between states can be proved using induction-based 
proof and fix-point calculation \cite{bjesse2000sat,stoffel1997record}.
\end{enumerate}

\begin{figure}[h]
\centering{
\includegraphics[scale=0.3]{newfig/fig_SAT.eps}
\caption{An example of equivalence checking on miter of circuit $A$ and $B$ using SAT}
\label{fig:miter}}
\end{figure}

A part of formal verification techniques adopt concepts and algorithms from 
\emph{computer-algebra} and \emph{algebraic geometry}.
Algebraic geometry provide a way to reason about the presence or absence of solutions 
without actually solving the system of constraints.
Using methods in \cite{Avrunin:CAV,condrat-tacas07,gbverify:2007,jinpeng,pruss:tcad15}, 
the circuit design can be transformed into a polynomial system. Subsequently, this system
of polynomials is canonicalized by computing a Gr\"obner basis (GB) \cite{gb_book}. 
Computation of GB allows for 
a straightforward proof of important properties of a polynomial system, 
such as the presence or absence of 
solutions. These properties can also be leveraged for 
verification. The disadvantage of the GB computation method is that its complexity can be doubly 
exponential in the worst case \cite{dube1986complexity}, and thus methods directly performing GB computation 
from an arbitrary setup is not practical for industry-level applications. However, recent
breakthroughs in computer-algebra hardware verification have shown
that it is possible to overcome the complexity of this computation while
still utilizing the beneficial properties of GB
\cite{lv:phd,tim:phd}.

Word-level abstractions have a wide variety of applications in formal 
verification. For example, it can work as automatic decision and canonical reduction engine in theorem proving;
for RTL composed of macro blocks, abstractions of these blocks are also benefit RTL verification.
Concretely, MC and equivalence checking with abstractions can be classified as:
\begin{itemize}  
\item Model checking with abstractions \cite{kroening:model}, 
where an over-pproximation of RTL blocks is abstracted and used to 
refine the RTL model.
\item Graph-based equivalence 
checking with abstraction \cite{WLS,arditi:bmd}, where abstraction 
generates a word-level canonical graph representation of the circuit.
\item SAT-based equivalence checking with abstraction \cite{lpsat}, where 
abstractions are used to analyze structural symmetries and similarities such that
the Boolean formulas fed to SAT solver is minimized.
\end{itemize}

Other equivalence checking techniques that employ abstractions 
include {\it satisfiability modulo theory} (SMT) solvers \cite{boolector,bryant:tacas07}.
as well as constraint programming (CP) techniques \cite{ms:research,tew:iccad08}.

Word-level abstractions also find applications in RTL and datapath 
synthesis \cite{demicheli:iccad_98,demicheli:dac_99,demicheli:tcad_03}. 
Since modern datapath design specifications are mostly word-level, synthesis tools with abstractions
can make use of larger macro blocks to generate and optimize the
datapaths. Moreover, 
word-level abstractions facilitates uninterpreted functions (UFs) \cite{UF3}, which 
can be transformed into proposition formulas with word-level information and verified using 
word-level theorem provers and model checkers.

\section{Abstractions in Sequential Design Verification}
With the increasing size of integrated
circuits, sequential circuit designers face complicated problems of
design errors in specification models and implementations. These
errors are usually modeled as ``bad" states, and the
circuits/functional components are modeled as finite state machines
(FSMs). Once state reachability is analyzed, the existence of errors
can be identified by checking whether ``bad" states are {\it
  reachable} from certain initial states. Temporal logic {\it model
  checking} (MC) formulations and solvers are often used for this
purpose. Once the designs and specification models are validated using
model checking, optimized implementations of sequential circuits are
synthesized. A subsequent problem then needs to be solved to prove
that the sequential circuit implementations are equivalent to the
original specification models, {\it i.e.} {\it Sequential Equivalence
  Checking }(SEC). When the specification is given as an arithmetic function
which canonically represents the circuit, then the problem 
becomes {\it functional verification} of sequential arithmetic circuits.

@@

For word-level SEC, we are given two designs, or their corresponding
Mealy/Moore FSMs ${\mathcal{M}}_1,{\mathcal{M}}_2$, along with their
initial starting states $S_0^1,S_0^2$. We wish to prove the absence of
a sequence of inputs (string) that distinguishes the initial
states \cite{coudert:iccad90}
\cite{coudert1990verification}. Fundamentally, this requires 
the construction of a product machine; and the main research problem
relates to that of performing {\it FSM traversal} \cite{touati1990implicit}
but  at the word-level. Analogously, in the case of MC, the problem
is setup {\it w.r.t.} a FSM $\mathcal{M}$, a set of initial states $S_0$ and
a set of property states $p$. Techniques are to be researched that
verify that there exist no sequence of transitions from an initial
state to a non-property state (``bad'' state). These problems have to
be solved in the context of word-level verification -- {\it i.e.} data
representation, abstraction using UNSAT cores and
algorithm execution has to be carried out at word-level.

@@

\section{Objective and Contribution of this Dissertation}
This research proposes a
set of new, promising approaches for {\it word-level representation,
reachability analysis and abstraction} for sequential design verification techniques. 
Our approaches operate at the word-level and are based
largely on the concepts from {\it algebraic geometry}.

@@

% TODO move to prev
% The use of algebraic geometry
% for sequential circuit verification and symbolic model checking has
% been presented before. Avrunin presented the
% concept of symbolic MC using algebraic geometry in
% \cite{Avrunin:CAV}. Later, in \cite{vardi-iasted07}, Vardi presented
% GB-algorithms for CTL, LTL and 
% bounded MC over Boolean rings. However, these approaches are a
% straight-forward transformation of the problem to {\it bit-level}
% Boolean GB engines which are used in lieu of BDDs or SAT solvers. All
% the concepts of word-level reachability, abstraction-refinement using
% interpolation or UNSAT cores, etc., that we desire was not the focus of
% \cite{Avrunin:CAV} \cite{vardi-iasted07}.

\subsection{Word-level Reachability Analysis}
In this dissertation, we propose a method to perform reachability analysis at the word-level. 
	The given FSM is modeled as a system of polynomials over a finite field,
	where the state space is mapped to the solutions of polynomial system.
	Our proposed algorithm utilizes concepts of ideal-variety correspondence in algebraic geometry.
	It also forms the foundation for word-level SEC and MC by enabling word-level abstraction of the
  state-space.

In this dissertation we represent
the FSMs -- the transition relations -- by means of a set of
multi-variate polynomials with coefficients from the finite (Galois)
field $\Fkk$ of $2^k$ elements, i.e. polynomials in the ring
$\Fkk[x_1,\dots,x_d]$. Each state of a FSM is identified with a
Boolean assignment to a set of $k$-bit state register variables
$S=\{s_0,\dots,s_{k-1}\}$. Therefore, we can consider each ($k$-bit)
state as a word-level element $S$ of the finite field
$\Fkk$. Algorithms can directly operate on polynomials in word-level
variable $S$.

Boolean functions with $k$-bit inputs and $k$-bit outputs 
$f: \B^k \rightarrow \B^k,\B = \{0, 1\}$ can be construed as functions
$f: \Fkk \rightarrow \Fkk$. It is well-known that over the finite
field ($\Fq$) of  $q$ elements, {\it every function} $f: \Fq
\rightarrow \Fq$ is a polynomial function \cite{ff:1997}. Moreover,
there exists a unique canonical polynomial $\Func$ that describes $f$.
This implies that one can derive a {\it  canonical, polynomial
  abstraction} of the function as $Z = \Func(A)$ where $Z, A$ are
word-level symbols representing $k$-bit operands. The concept also
generalizes to functions with different input/output bit-vector sizes,
i.e. functions of the type $f: \B^n \rightarrow \B^m$, modeled as
polynomials over $f:{\mathbb{F}}_{2^k} \rightarrow
{\mathbb{F}}_{2^k}$, where $k=LCM(n,m)$ \cite{ff:1997}.  
This implies that the FSM's transition relations can be
represented as polynomial functions (ideals) in $\Fkk$, and values of
state variables can be represented as solutions to these polynomials
(variety of the ideal). Subsequently, the ideal-variety correspondences
in algebraic geometry can be applied to implement symbolic reasoning
about state reachability. 
% Moreover, as ${\mathbb{F}}_2 \subset \Fkk$,
% our model provides a single, unified and bit-precise representation
% for both bit-level (${\mathbb{F}}_2$) and word-level ($\Fkk$)
% constraints.

The decision and abstraction procedures in our setting will rely on
the theory and technology of {\it \Grobner bases}. GB-based
algebraic reasoning is very powerful; in fact it is known to be
strictly stronger than resolution \cite{CEI:stoc-96}. Therefore, in
light of the above discussion, using concepts from algebraic geometry
and \Grobner bases over $\Fkk$, we can introduce another dimension of
word-level abstraction to the techniques in sequential verification.

\subsection{Application to Sequential Galois Field Arithmetic Circuits Verification}
Sequential Galois field (GF) arithmetic circuits can be modeled as a special type of 
FSM, where the pre-loaded operands are mapped to initial states and pseudo outputs
after $k$ clock-cycles are recognized as the final reached state after $k$ transitions.
Therefore, the word-level FSM traversal algorithm can be applied to 
verify the correctness of final reached state, {\it i.e.} the functional correctness of
a sequential arithmetic circuit.

In our proposed approach, word-level abstraction is employed to generate the 
word-level signature of the combinational logic component of the sequential 
arithmetic circuit. This abstraction requires GB computation, which usually has high time/space 
complexity. We propose several improvements to simplify the GB computation procedure
and make the entire algorithm execution practical. As a result, we successfully verify sequential 
multipliers with 162-bit datapaths.

\subsection{UNSAT Cores in Algebraic Geometry for Abstraction Refinement}
Abstraction is an effective method to lower the cost to traverse the state space.
In modern model checkers, abstraction is used to refine the model 
during the iterative execution of the tool. UNSAT core is widely used 
as an important component of abstraction refinement. The reason is that UNSAT cores can provide 
information to indicate the redundancy in the state variables, and that information is necessary for the refinement 
process.

In this dissertation, we explore the concept, and the computation, of unsatisfiable (UNSAT)
  cores of  a set of polynomials using Gr\"obner bases calculation. We also propose
a number of heuristics by cracking the Buchberger's algorithm to reduce the size of UNSAT core. 
  We apply the UNSAT core extraction 
  to a bounded model checker (BMC) with abstraction refinement.

% \subsection{Dissertation Contributions}
% In this dissertation, we focus on sequential circuit verification problems and 
% propose solutions consisting of three main contributions: 
% \begin{enumerate}[{1)}]
% \item A method to perform reachability analysis at the word-level. 
% 	The given FSM is modeled as a system of polynomials over finite field,
% 	where the state space is mapped to its solution space.
% 	Our proposed algorithm utilize concepts in algebraic geometry including ideals and varieties.
% 	It also forms the foundation for word-level SEC and MC by enabling word-level abstraction of the
%   state-space.
% \item Using the theory of FSM traversal, we apply the algorithm to verify the 
% function of sequential GF arithmetic circuits. Our proposed approach uses GB-engines efficiently 
% and can verify sequential multipliers with 162-bit datapaths.
% \item We explore the concept, and the computation, of unsatisfiable (UNSAT)
%   cores of  a set of polynomials using Gr\"obner bases calculation. We apply the UNSAT core extraction 
%   to abstraction-refinement
%   techniques such as bounded model checking (BMC). 
% \end{enumerate}

\section{Dissertation Organization}
The rest of the dissertation is organized as
follows: Chapter \ref{ch:prev} reviews previous work, and analyzes their drawbacks with respect to 
the word-level sequential verification problem.
Chapter \ref{ch:prelim_GF} covers preliminary
concepts and notation on finite fields, and the methodology about design of arithmetic circuits in finite fields.
Chapter \ref{ch:ideals} provides a theoretical background on algebraic geometry and Gr\"obner bases.
Chapter \ref{ch:reacha} describes the basic
concept of word-level FSM traversal and introduces our proposed word-level FSM traversal algorithm. 
Chapter \ref{ch:normal} explores the application of FSM traversal algorithm on 
functional verification of sequential GF arithmetic circuits. Chapter \ref{ch:UNSAT} describes
algorithmic techniques to derive UNSAT cores of polynomial
ideals. It also demonstrates with the help of an example how abstraction via
UNSAT cores in algebraic geometry can simplify BMC. 
Chapter \ref{ch:conclude} concludes the dissertation and outlines potential future research for
continuation of this work. Appendix provides theory and methodology on characterization of normal basis,
as well as the construction of optimal normal basis and application to normal basis multiplier design.
